---
title: "Getting started with readAOFM"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with readAOFM}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)

fixture_path <- function(filename) {
  installed <- system.file("extdata", filename, package = "readAOFM")
  if (nzchar(installed) && file.exists(installed)) {
    return(installed)
  }

  source_candidates <- c(
    file.path("inst", "extdata", filename),
    file.path("..", "inst", "extdata", filename)
  )
  source <- source_candidates[file.exists(source_candidates)]
  if (length(source) > 0L) {
    return(normalizePath(source[[1]], mustWork = TRUE))
  }

  stop(sprintf("Packaged fixture '%s' was not found.", filename), call. = FALSE)
}

tb_issuance_fixture <- fixture_path("tb_issuance.xlsx")
```

`readAOFM` reads workbook data published by the Australian Office of Financial
Management (AOFM) Data Hub and returns R-friendly data frames or named lists.
The examples in this article separate the two parts of a normal workflow:
discover a table locally, then read a selected workbook. The first successful
read uses a packaged workbook snapshot, so it is deterministic and does not
contact AOFM.

## Find a table without downloading anything

`search_aofm()` searches the package's local catalogue. By default it returns
metadata and ready-to-run calls; it does not make a network request.

```{r search-first}
readAOFM::search_aofm("tb issuance")[, c("id", "reader", "read_call")]
```

The search accepts ordinary descriptions as well as exact terminology. For
example, these searches identify the indexed-bond and non-resident ownership
families:

```{r search-language}
readAOFM::search_aofm("inflation")
readAOFM::search_aofm("foreign ownership")
```

The `read_call` column is useful when a query returns more than one table. For
example, `search_aofm("issuance")` finds Treasury Bond, Treasury Indexed Bond,
and Treasury Note issuance; reading all three would perform three workbook
downloads.

## Read the first table using a local fixture

The public entry point is `read_aofm(security, type, csv = FALSE)`. The
`security` and `type` values are exact selectors. A single match returns that
table's parsed object. The package's tests and vignettes use
`testthat::with_mocked_bindings()` only to replace the internal downloader with
the packaged workbook; the call below still exercises the public dispatcher
and the real parser.

```{r first-read}
tb_issuance <- testthat::with_mocked_bindings(
  readAOFM::read_aofm("tb", "issuance"),
  download_aofm_table_workbook = function(aofm_table) {
    stopifnot(identical(aofm_table, "tb_issuance"))
    tb_issuance_fixture
  },
  .package = "readAOFM"
)

stopifnot(
  is.data.frame(tb_issuance),
  all(c("date_held", "tender_number", "maturity", "name", "value") %in%
        names(tb_issuance)),
  inherits(tb_issuance$date_held, "Date")
)

data.frame(
  object_class = paste(class(tb_issuance), collapse = ", "),
  rows = nrow(tb_issuance),
  columns = ncol(tb_issuance),
  date_class = paste(class(tb_issuance$date_held), collapse = ", "),
  stringsAsFactors = FALSE
)
```

The transactional reader preserves workbook identifiers and pivots numeric
measures into the long-form `name`/`value` columns. Date fields handled by the
reader are returned as `Date` objects. Exact source columns can change if AOFM
changes its workbook, so inspect names before building a downstream analysis.

```{r first-read-preview}
utils::head(tb_issuance[c("date_held", "maturity", "name", "value")], 3)
```

For a single table, `read_aofm()` returns the parsed data frame directly. When
selectors match several supported tables, it returns a named list keyed by
table ID; omitting both selectors requests all 23 parser-supported tables and
can therefore involve many downloads. `csv = TRUE` keeps the same return value
and additionally writes the reader's CSV output under `output/` in the current
working directory; see the network and outputs article for an isolated example.

## Optional live read

The normal live call is intentionally guarded. Set
`READAOFM_RUN_LIVE_EXAMPLES=true` in an interactive or development session to
opt in after checking that the AOFM site is reachable. The default vignette
build path never makes a network request.

```{r live-read}
if (interactive() && identical(Sys.getenv("READAOFM_RUN_LIVE_EXAMPLES"), "true")) {
  live_tb_issuance <- readAOFM::read_aofm("tb", "issuance")
  stopifnot(is.data.frame(live_tb_issuance))
  utils::head(live_tb_issuance, 3)
} else {
  cat("Live read skipped; set READAOFM_RUN_LIVE_EXAMPLES=true to opt in.\n")
}
```

Live readers fetch public workbooks over HTTPS and do not require AOFM
credentials. They stage each workbook in a temporary file rather than using a
persistent package download cache. Upstream availability and workbook layout
remain outside the package's control.
