---
title: "Getting started with readAOFM"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with readAOFM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteDepends{ggplot2,testthat}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)

fixture_path <- function(filename) {
  installed <- system.file("extdata", filename, package = "readAOFM")
  if (nzchar(installed) && file.exists(installed)) {
    return(installed)
  }

  source_candidates <- c(
    file.path("inst", "extdata", filename),
    file.path("..", "inst", "extdata", filename)
  )
  source <- source_candidates[file.exists(source_candidates)]
  if (length(source) > 0L) {
    return(normalizePath(source[[1]], mustWork = TRUE))
  }

  stop(sprintf("Packaged fixture '%s' was not found.", filename), call. = FALSE)
}

tb_issuance_fixture <- fixture_path("tb_issuance.xlsx")

# The article executes the public dispatcher and production parser while
# substituting the installed official snapshot at the transport boundary.
tb_issuance <- suppressMessages(testthat::with_mocked_bindings(
  readAOFM::read_aofm("tb", "issuance"),
  download_aofm_table_workbook = function(aofm_table) {
    stopifnot(identical(aofm_table, "tb_issuance"))
    tb_issuance_fixture
  },
  .package = "readAOFM"
))
```

`readAOFM` connects local catalogue discovery with retrieval and parsing of
workbooks published by the Australian Office of Financial Management (AOFM)
Data Hub. This article follows a complete workflow: identify Treasury Bond
issuance, retrieve its long-form observations, select a representative series,
and create a chart. A packaged official snapshot makes the rendered article
deterministic while the user-facing call remains the normal public API.

## Discover a table locally

`search_aofm()` searches the package's local catalogue and returns metadata and
ready-to-run calls. The default `read = FALSE` keeps discovery network-free.

```{r search-first}
readAOFM::search_aofm("tb issuance")[, c("id", "reader", "read_call")]
```

The search accepts ordinary descriptions as well as exact terminology. For
example, these searches identify the indexed-bond and non-resident ownership
families:

```{r search-language}
readAOFM::search_aofm("inflation")
readAOFM::search_aofm("foreign ownership")
```

The `read_call` column is useful when a query returns more than one table. For
example, `search_aofm("issuance")` finds Treasury Bond, Treasury Indexed Bond,
and Treasury Note issuance; retrieving that full result would request three
workbooks.

## Retrieve and chart a time series

The public entry point is `read_aofm(security, type, csv = FALSE)`. The
`security` and `type` values are exact selectors. For the current AOFM source,
run:

```{r retrieve-current, eval=FALSE}
tb_issuance <- readAOFM::read_aofm("tb", "issuance")
```

The rendered article executes that same public dispatcher and parser with the
installed `tb_issuance.xlsx` snapshot substituted at the HTTPS transport
boundary. The snapshot comes from
[AOFM media/591](https://www.aofm.gov.au/media/591) and has SHA-256
`4f74568d37258a6fad7b80136cdd64a29341f7bfa8550d2a4d7f8cc785e2e5c9`.
The packaged snapshot documentation records its provenance, attribution, and
licensing context. Live calls retrieve the current workbook instead.

The public return contract is visible directly from the parsed object:

```{r parsed-contract}
stopifnot(
  is.data.frame(tb_issuance),
  all(c("date_held", "tender_number", "maturity", "name", "value") %in%
        names(tb_issuance)),
  inherits(tb_issuance$date_held, "Date"),
  inherits(tb_issuance$maturity, "Date"),
  is.numeric(tb_issuance$value)
)

data.frame(
  object_class = paste(class(tb_issuance), collapse = ", "),
  rows = nrow(tb_issuance),
  columns = ncol(tb_issuance),
  date_class = paste(class(tb_issuance$date_held), collapse = ", "),
  stringsAsFactors = FALSE
)
```

The transactional reader preserves workbook identifiers, represents measures
in the long-form `name`/`value` columns, and returns handled dates as `Date`
objects. The following base-R transformation selects two source measures for
the 21 April 2029 maturity and converts their dollar values to A$ billions.

```{r prepare-time-series}
measure_labels <- c(
  amount_allotted = "Amount allotted",
  amount_of_bids = "Bids received"
)

tender_series <- tb_issuance[
  tb_issuance$maturity == as.Date("2029-04-21") &
    tb_issuance$name %in% names(measure_labels) &
    !is.na(tb_issuance$value),
  c("date_held", "maturity", "name", "value")
]

tender_series$measure <- factor(
  tender_series$name,
  levels = names(measure_labels),
  labels = unname(measure_labels)
)
tender_series$value_billions <- tender_series$value / 1e9
tender_series <- tender_series[
  order(tender_series$measure, tender_series$date_held),
  ,
  drop = FALSE
]

series_preview <- utils::head(tender_series, 4)
data.frame(
  date_held = series_preview$date_held,
  measure = as.character(series_preview$measure),
  value_billions = sprintf("%.2f", series_preview$value_billions),
  stringsAsFactors = FALSE
)
```

```{r retrieve-chart, fig.width=8, fig.height=4.8, dpi=144, fig.retina=1, out.width="100%", fig.alt="Line chart of bids received and amounts allotted at Australian Treasury Bond tenders for the 21 April 2029 maturity, from 2012 to 2025."}
if (requireNamespace("ggplot2", quietly = TRUE)) {
  tender_plot <- ggplot2::ggplot(
    tender_series,
    ggplot2::aes(
      x = date_held,
      y = value_billions,
      colour = measure,
      group = measure
    )
  ) +
    ggplot2::geom_line(linewidth = 0.75) +
    ggplot2::geom_point(size = 1.25) +
    ggplot2::scale_colour_manual(values = c("#0072B2", "#D55E00")) +
    ggplot2::labs(
      title = "Treasury Bond tender bids and allotments",
      subtitle = "21 April 2029 maturity, packaged AOFM snapshot",
      x = "Tender date",
      y = "A$ billions",
      colour = NULL,
      caption = paste(
        "Source: Australian Office of Financial Management (AOFM),",
        "Data Hub."
      )
    ) +
    ggplot2::theme_minimal(base_size = 11) +
    ggplot2::theme(
      legend.position = "bottom",
      legend.justification = "left",
      panel.grid.minor = ggplot2::element_blank(),
      plot.caption = ggplot2::element_text(hjust = 0, size = 8)
    )

  tender_plot
}
```

For a single table, `read_aofm()` returns the parsed data frame directly. When
selectors match several supported tables, it returns a named list keyed by
table ID. Omitting both selectors requests all 23 parser-supported tables and
can involve many downloads. `csv = TRUE` preserves the parsed return value and
also writes the reader's CSV output under `output/` in the current working
directory; the network and outputs article provides an isolated example.

## Retrieve the latest AOFM workbook

Set `READAOFM_RUN_LIVE_EXAMPLES=true` in an interactive or development session
to run the live call below after checking that the AOFM site is reachable. The
ordinary article build uses the deterministic snapshot path above.

```{r live-read}
if (interactive() && identical(Sys.getenv("READAOFM_RUN_LIVE_EXAMPLES"), "true")) {
  live_tb_issuance <- readAOFM::read_aofm("tb", "issuance")
  stopifnot(is.data.frame(live_tb_issuance))
  utils::head(live_tb_issuance, 3)
} else {
  cat("Live read skipped; set READAOFM_RUN_LIVE_EXAMPLES=true to opt in.\n")
}
```

Live readers fetch public workbooks over HTTPS without AOFM credentials. They
stage each workbook in a temporary file and retrieve the current source on
every call. AOFM controls upstream availability and workbook layout; use the
network, outputs, and troubleshooting article for bounded transport details
and recovery steps.
