---
title: "Getting started with readAOFM"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with readAOFM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteDepends{ggplot2,testthat}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)

fixture_path <- function(filename) {
  installed <- system.file("extdata", filename, package = "readAOFM")
  if (nzchar(installed) && file.exists(installed)) {
    return(installed)
  }

  source_candidates <- c(
    file.path("inst", "extdata", filename),
    file.path("..", "inst", "extdata", filename)
  )
  source <- source_candidates[file.exists(source_candidates)]
  if (length(source) > 0L) {
    return(normalizePath(source[[1]], mustWork = TRUE))
  }

  stop(sprintf("Packaged fixture '%s' was not found.", filename), call. = FALSE)
}

tb_issuance_fixture <- fixture_path("tb_issuance.xlsx")

# The article executes the public dispatcher and production parser while
# substituting the installed official snapshot at the transport boundary.
tb_issuance <- suppressMessages(testthat::with_mocked_bindings(
  readAOFM::read_aofm("tb", "issuance"),
  download_aofm_table_workbook = function(aofm_table, ...) {
    stopifnot(identical(aofm_table, "tb_issuance"))
    tb_issuance_fixture
  },
  .package = "readAOFM"
))
```

`readAOFM` connects local catalogue discovery with retrieval and parsing of
workbooks published by the Australian Office of Financial Management (AOFM)
Data Hub. This article follows a complete workflow: identify Treasury Bond
issuance, retrieve its long-form observations, select a representative series,
and create a chart. A packaged official snapshot makes the rendered article
deterministic while the user-facing call remains the normal public API.

## Discover a table locally

`search_aofm()` searches the package's local catalogue and returns metadata and
ready-to-run calls. The default `read = FALSE` keeps discovery network-free.

```{r search-first}
readAOFM::search_aofm("tb issuance")[, c("id", "reader", "read_call")]
```

The search accepts ordinary descriptions as well as exact terminology. For
example, these searches identify the indexed-bond and non-resident ownership
families:

```{r search-language}
readAOFM::search_aofm("inflation")
readAOFM::search_aofm("foreign ownership")
```

The `read_call` column is useful when a query returns more than one table. For
example, `search_aofm("issuance")` finds Treasury Bond, Treasury Indexed Bond,
and Treasury Note issuance; retrieving that full result would request three
workbooks.

## Retrieve and chart a time series

The public entry point is `read_aofm(security, type, csv = FALSE)`. The
`security` and `type` values are exact selectors. For the current AOFM source,
run:

```{r retrieve-current, eval=FALSE}
tb_issuance <- readAOFM::read_aofm("tb", "issuance")
```

The rendered article executes that same public dispatcher and parser with the
installed `tb_issuance.xlsx` snapshot substituted at the HTTPS transport
boundary. The snapshot was originally acquired as AOFM media item 591; the
packaged snapshot documentation records that historical acquisition, its
verified SHA-256, attribution, and licensing context. That identifier is not a
live route. Ordinary calls retrieve the current direct HTTPS Data Hub workbook
route from the local catalogue.

The public return contract is visible directly from the parsed object:

```{r parsed-contract}
stopifnot(
  is.data.frame(tb_issuance),
  all(c("date_held", "tender_number", "maturity", "name", "value") %in%
        names(tb_issuance)),
  inherits(tb_issuance$date_held, "Date"),
  inherits(tb_issuance$maturity, "Date"),
  is.numeric(tb_issuance$value)
)

data.frame(
  object_class = paste(class(tb_issuance), collapse = ", "),
  rows = nrow(tb_issuance),
  columns = ncol(tb_issuance),
  date_class = paste(class(tb_issuance$date_held), collapse = ", "),
  stringsAsFactors = FALSE
)
```

The transactional reader preserves workbook identifiers, represents measures
in the long-form `name`/`value` columns, and returns handled dates as `Date`
objects. The following base-R transformation selects two source measures for
the 21 April 2029 maturity and converts their dollar values to A$ billions.

```{r prepare-time-series}
series_labels <- c(
  amount_allotted = "Amount allotted",
  amount_of_bids = "Bids received"
)

selected <- tb_issuance[
  tb_issuance$maturity == as.Date("2029-04-21") &
    tb_issuance$name %in% names(series_labels) &
    !is.na(tb_issuance$value),
  c("date_held", "name", "value")
]

chart_data <- data.frame(
  date = selected$date_held,
  value = selected$value / 1e9,
  series = unname(series_labels[selected$name]),
  stringsAsFactors = FALSE
)
chart_data <- chart_data[order(chart_data$series, chart_data$date), ]
utils::head(chart_data, 4)
```

```{r retrieve-chart, fig.width=8, fig.height=4.8, dpi=144, fig.retina=1, out.width="100%", fig.alt="Line chart of bids received and amounts allotted at Australian Treasury Bond tenders for the 21 April 2029 maturity, from 2012 to 2025."}
if (requireNamespace("ggplot2", quietly = TRUE)) {
  ggplot2::ggplot(
    chart_data,
    ggplot2::aes(x = date, y = value, colour = series, group = series)
  ) +
    ggplot2::geom_line() +
    ggplot2::labs(
      title = "Treasury Bond tender bids and allotments",
      subtitle = "Published tender observations for the 21 April 2029 maturity",
      x = "Tender date",
      y = "A$ billions",
      colour = "Measure",
      caption = "Source: Australian Office of Financial Management Data Hub"
    ) +
    ggplot2::theme_minimal(base_size = 13)
}
```

For a single table, `read_aofm()` returns the parsed data frame directly. When
selectors match several supported tables, it returns a named list keyed by
table ID. Omitting both selectors requests all 23 parser-supported tables and
can involve many downloads. `csv = TRUE` preserves the parsed return value and
also writes the reader's CSV output under `output/` in the current working
directory; the network and outputs article provides an isolated example.

The runtime catalogue tracks 30 IDs in total: 23 parser-supported tables and
seven raw-only/unsupported rows. Use `aofm_catalog(include_unsupported = TRUE)`
to inspect all direct HTTPS source routes and filename extensions before a raw
download. `portfolio_overview` remains a raw-only historical direct workbook
because it is not currently listed on the Data Hub; no parser support is
implied.

## Turnover source transition

The two turnover readers stitch the historical and redesigned current AOFM
workbooks. Historical turnover covers July 2016 through December 2025; the
redesigned current workbooks begin in January 2026, with no gap at the
boundary. Current workbooks contain monthly observation periods, while the Data
Hub is updated quarterly with an approximately two-month lag. Historical
turnover has mixed source granularity: **By Tenor** is monthly and **By
Category** is quarterly. Current sheets provide `security`, `region`, and
`counterparty`; the combined result uses `tenor`, `investor_type`, `security`,
`region`, and `counterparty` groups.

Rows are identified by the natural key `period`, `group`, and `name`. Current
rows take precedence if the two sources overlap, and duplicate keys within a
source are rejected. The result carries a two-record `aofm_sources` attribute
with historical/current roles and raw URL, filename, byte-count, and SHA-256
metadata. The durable route and boundary record is in
`inst/extdata/README-aofm-route-continuity.md`.

## Retrieve the latest AOFM workbook

Set `READAOFM_RUN_LIVE_EXAMPLES=true` in an interactive or development session
to run the live call below after checking that the AOFM site is reachable. The
ordinary article build uses the deterministic snapshot path above.

```{r live-read}
if (interactive() && identical(Sys.getenv("READAOFM_RUN_LIVE_EXAMPLES"), "true")) {
  live_tb_issuance <- readAOFM::read_aofm("tb", "issuance")
  stopifnot(is.data.frame(live_tb_issuance))
  utils::head(live_tb_issuance, 3)
} else {
  cat("Live read skipped; set READAOFM_RUN_LIVE_EXAMPLES=true to opt in.\n")
}
```

Live readers fetch public workbooks over HTTPS without AOFM credentials. They
stage each workbook in a temporary file and retrieve the current source on
every call. AOFM controls upstream availability and workbook layout; use the
network, outputs, and troubleshooting article for bounded transport details
and recovery steps. The public transport controls are `timeout` (30 seconds
per attempt by default), `retries` (one retry by default), and `max_bytes` (100
MiB by default).

For an explicit caller-managed copy, `download_aofm_file()` stores verified
SHA-256-named workbooks and metadata below `.readAOFM/data/<table-id>/` beneath
the chosen root (by default `tempdir()`), with bounded pruning and writer
locking. It is separate from
temporary reader staging and the legacy `data/`/`output/` writes. Use
`aofm_file_metadata()` to inspect a local digest and `read_aofm_file()` to
parse a local workbook without another network request.
