---
title: "Network, outputs, and troubleshooting"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Network, outputs, and troubleshooting}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)

fixture_path <- function(filename) {
  installed <- system.file("extdata", filename, package = "readAOFM")
  if (nzchar(installed) && file.exists(installed)) {
    return(installed)
  }

  source_candidates <- c(
    file.path("inst", "extdata", filename),
    file.path("..", "inst", "extdata", filename)
  )
  source <- source_candidates[file.exists(source_candidates)]
  if (length(source) > 0L) {
    return(normalizePath(source[[1]], mustWork = TRUE))
  }

  stop(sprintf("Packaged fixture '%s' was not found.", filename), call. = FALSE)
}

tb_issuance_fixture <- fixture_path("tb_issuance.xlsx")

with_temp_working_directory <- function(code) {
  old <- getwd()
  work <- tempfile("readAOFM-vignette-")
  dir.create(work, recursive = TRUE)
  on.exit(setwd(old), add = TRUE)
  setwd(work)
  force(code)
}
```

`readAOFM` separates local catalogue discovery from workbook I/O. Readers fetch
public AOFM Data Hub workbooks and can optionally write parsed CSV files, while
`search_aofm()` searches installed metadata. The examples below keep every
write in an isolated temporary directory and use a packaged workbook snapshot
for deterministic parsing. The package's managed download API is explicit, so
ordinary reader calls do not create a persistent cache.

## Network and staging behavior

The reader path resolves a table from the local catalogue, requests its AOFM
workbook over HTTPS, validates that the response looks like an Excel workbook,
and stages the file in a temporary location for parsing. Public workbooks are
available without package credentials, and each live read retrieves the
current source. Transport failures, HTTP errors, empty files, non-workbook
responses, and changed source layouts are reported with table context.

Use the offline search before a live call:

```{r offline-search}
readAOFM::search_aofm("tb issuance")[, c("id", "read_call")]
```

## CSV output is opt-in

`csv = TRUE` preserves the parsed return value and asks the family reader to
write a CSV beneath `output/`. The following example uses the real public
`read_aofm()` and parser, but mocks only the internal downloader so that the
result is repeatable and no network request is made.

```{r csv-side-effect}
csv_probe <- with_temp_working_directory({
  parsed <- testthat::with_mocked_bindings(
    readAOFM::read_aofm("tb", "issuance", csv = TRUE),
    download_aofm_table_workbook = function(aofm_table, ...) {
      stopifnot(identical(aofm_table, "tb_issuance"))
      tb_issuance_fixture
    },
    .package = "readAOFM"
  )

  output_file <- file.path("output", "tb_issuance.csv")
  stopifnot(file.exists(output_file))

  list(
    parsed = parsed,
    files = list.files(".", recursive = TRUE, all.files = FALSE),
    csv = data.frame(
      path = output_file,
      bytes = unname(file.info(output_file)$size),
      stringsAsFactors = FALSE
    )
  )
})

csv_probe$files
csv_probe$csv
utils::head(csv_probe$parsed[c("date_held", "name", "value")], 3)
```

Leaving `csv = FALSE` (the default) does not create an `output/` directory.
The raw helper `download_aofm_xlsx()` has different behavior: it saves source
workbooks beneath `data/` and is intended for an explicitly writable project
directory. Neither output path is a package cache. The public transport bounds
are `timeout` (30 seconds per attempt by default), `retries` (one retry), and
`max_bytes` (100 MiB); pass them explicitly when a script needs tighter
limits.

## Managed downloads and offline parsing

`download_aofm_file()` is the opt-in workflow for retaining a verified current
workbook. It stores a SHA-256-named `.xls`/`.xlsx` file, `current.rds`
provenance, and a short-lived writer lock beneath the caller's
`.readAOFM/data/<table-id>/` directory. Age, file-count, and total-byte limits
prune old entries. The cache root defaults to `tempdir()` and can be set with
the function's `path` argument. The returned path carries source URL, filename,
byte-count, SHA-256, retrieval time, and cache-hit metadata.

`aofm_file_metadata()` computes local byte-level provenance without network or
cache writes. `read_aofm_file()` parses a local workbook using the same family
parser selected by its stable table ID. This is useful for a downloaded file or
an installed fixture:

```{r local-file-probe}
local_metadata <- readAOFM::aofm_file_metadata(
  tb_issuance_fixture,
  table_id = "tb_issuance"
)
local_parsed <- readAOFM::read_aofm_file(
  tb_issuance_fixture,
  table_id = "tb_issuance"
)
local_metadata[c("table_id", "raw_bytes", "raw_sha256")]
utils::head(local_parsed[c("date_held", "name", "value")], 3)
```

The local path is not silently treated as a current source: standalone
metadata records its local modification time and omits a source URL. Managed
cache metadata remains qualified only while its file, digest, and catalogue
identity agree.

## Multi-result calls and cost

Selectors can intentionally identify several tables. For example,
`read_aofm(type = "issuance")` returns a named list for Treasury Bond, Treasury
Indexed Bond, and Treasury Note issuance, and it downloads each workbook. A
security-only call such as `read_aofm("tb")` similarly reads every supported
Treasury Bond table. Search first to see the exact set and generated calls:

```{r multi-result}
issuance <- readAOFM::search_aofm("issuance")
issuance[, c("security", "type", "id", "read_call")]
```

Start with a single `security`/`type` pair when you are exploring, then make a
multi-result call once the additional network and parsing cost is intended.

## Common errors and recovery

Selector validation happens locally before any AOFM request:

```{r invalid-selector}
tryCatch(
  readAOFM::read_aofm("not-a-security"),
  error = function(error) conditionMessage(error)
)
```

For a live download or parse failure, check the following in order:

1. Run `search_aofm()` and copy the generated `read_call` so that selectors
   are exact.
2. Check network/proxy/SSL access to the AOFM source URL and retry once.
3. If the error mentions HTML, HTTP status, or workbook signature, AOFM may
   have returned an error page or redirect instead of an Excel workbook.
4. If the error mentions sheets or required columns, the upstream workbook
   layout may have changed. Report the table ID, full error, and access date.
5. If a CSV or raw download cannot be written, switch to a writable project
   directory or leave `csv = FALSE`.

For deterministic checks, use the packaged snapshots with
`read_aofm_file()` as above and keep any temporary working directory isolated
from analysis outputs. The legacy readers intentionally remain catalogue-first;
the local-file API separates retrieval from parsing when a caller needs that
control.

## Turnover source transition

`read_secondary("tb_turnover")` and `read_secondary("tib_turnover")` stitch
historical companions covering July 2016 through December 2025 to redesigned
current workbooks beginning in January 2026. Current workbooks contain monthly
observation periods, while the Data Hub is updated quarterly with an
approximately two-month lag. Historical **By Tenor** observations are monthly;
historical **By Category** observations are quarterly. Current sheets provide
`security`, `region`, and `counterparty`; the combined groups are `tenor`,
`investor_type`, `security`, `region`, and `counterparty`.

The natural key is `period`, `group`, and `name`. Current rows take precedence
on overlap, duplicate keys within a source are rejected, and the returned data
frame carries a two-record `aofm_sources` attribute with historical/current
roles and raw SHA-256 metadata. See the route-continuity record for the exact
source URLs and boundary evidence.

## Optional live troubleshooting probe

This normal live call is opt-in. It is disabled during ordinary vignette and
CRAN builds, so rendering remains deterministic and offline.

```{r live-troubleshooting}
if (interactive() && identical(Sys.getenv("READAOFM_RUN_LIVE_EXAMPLES"), "true")) {
  live <- readAOFM::read_aofm("tb", "issuance")
  utils::head(live, 3)
} else {
  cat("Live probe skipped; set READAOFM_RUN_LIVE_EXAMPLES=true to opt in.\n")
}
```
