---
title: "Network, outputs, and troubleshooting"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Network, outputs, and troubleshooting}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)

fixture_path <- function(filename) {
  installed <- system.file("extdata", filename, package = "readAOFM")
  if (nzchar(installed) && file.exists(installed)) {
    return(installed)
  }

  source_candidates <- c(
    file.path("inst", "extdata", filename),
    file.path("..", "inst", "extdata", filename)
  )
  source <- source_candidates[file.exists(source_candidates)]
  if (length(source) > 0L) {
    return(normalizePath(source[[1]], mustWork = TRUE))
  }

  stop(sprintf("Packaged fixture '%s' was not found.", filename), call. = FALSE)
}

tb_issuance_fixture <- fixture_path("tb_issuance.xlsx")

with_temp_working_directory <- function(code) {
  old <- getwd()
  work <- tempfile("readAOFM-vignette-")
  dir.create(work, recursive = TRUE)
  on.exit(setwd(old), add = TRUE)
  setwd(work)
  force(code)
}
```

`readAOFM` has two kinds of I/O: readers fetch public AOFM Data Hub workbooks and
optionally write parsed CSV files; `search_aofm()` is local and has no I/O. The
examples below keep all writes in an isolated temporary directory and replace
the network downloader with a packaged workbook snapshot.

## Network and staging behavior

The reader path resolves a table from the local catalogue, requests its AOFM
workbook over HTTPS, validates that the response looks like an Excel workbook,
and stages the file in a temporary location for parsing. It does not require
credentials and does not maintain a persistent package cache. Repeating a live
read can therefore observe a newer AOFM workbook. Transport failures, HTTP
errors, empty files, non-workbook responses, and changed source layouts are
reported with table context.

Use the offline search before a live call:

```{r offline-search}
readAOFM::search_aofm("tb issuance")[, c("id", "read_call")]
```

## CSV output is opt-in

`csv = TRUE` preserves the parsed return value and asks the family reader to
write a CSV beneath `output/`. The following example uses the real public
`read_aofm()` and parser, but mocks only the internal downloader so that the
result is repeatable and no network request is made.

```{r csv-side-effect}
csv_probe <- with_temp_working_directory({
  parsed <- testthat::with_mocked_bindings(
    readAOFM::read_aofm("tb", "issuance", csv = TRUE),
    download_aofm_table_workbook = function(aofm_table) {
      stopifnot(identical(aofm_table, "tb_issuance"))
      tb_issuance_fixture
    },
    .package = "readAOFM"
  )

  output_file <- file.path("output", "tb_issuance.csv")
  stopifnot(file.exists(output_file))

  list(
    parsed = parsed,
    files = list.files(".", recursive = TRUE, all.files = FALSE),
    csv = data.frame(
      path = output_file,
      bytes = unname(file.info(output_file)$size),
      stringsAsFactors = FALSE
    )
  )
})

csv_probe$files
csv_probe$csv
utils::head(csv_probe$parsed[c("date_held", "name", "value")], 3)
```

Leaving `csv = FALSE` (the default) does not create an `output/` directory.
The raw helper `download_aofm_xlsx()` has different behavior: it saves source
workbooks beneath `data/` and is intended for an explicitly writable project
directory. Neither output path is a package cache.

## Multi-result calls and cost

Selectors can intentionally identify several tables. For example,
`read_aofm(type = "issuance")` returns a named list for Treasury Bond, Treasury
Indexed Bond, and Treasury Note issuance, and it downloads each workbook. A
security-only call such as `read_aofm("tb")` similarly reads every supported
Treasury Bond table. Search first to see the exact set and generated calls:

```{r multi-result}
issuance <- readAOFM::search_aofm("issuance")
issuance[, c("security", "type", "id", "read_call")]
```

Start with a single `security`/`type` pair when you are exploring, then make a
multi-result call once the additional network and parsing cost is intended.

## Common errors and recovery

An invalid selector is detected locally and does not contact AOFM:

```{r invalid-selector}
tryCatch(
  readAOFM::read_aofm("not-a-security"),
  error = function(error) conditionMessage(error)
)
```

For a live download or parse failure, check the following in order:

1. Run `search_aofm()` and copy the generated `read_call` so that selectors
   are exact.
2. Check network/proxy/SSL access to the AOFM source URL and retry once.
3. If the error mentions HTML, HTTP status, or workbook signature, AOFM may
   have returned an error page or redirect instead of an Excel workbook.
4. If the error mentions sheets or required columns, the upstream workbook
   layout may have changed. Report the table ID, full error, and access date.
5. If a CSV or raw download cannot be written, switch to a writable project
   directory or leave `csv = FALSE`.

The package does not accept a local workbook path through the public readers.
For deterministic checks, use the packaged snapshots as above and keep any
temporary working directory isolated from analysis outputs.

## Optional live troubleshooting probe

This normal live call is opt-in. It is disabled during ordinary vignette and
CRAN builds, so rendering remains deterministic and offline.

```{r live-troubleshooting}
if (interactive() && identical(Sys.getenv("READAOFM_RUN_LIVE_EXAMPLES"), "true")) {
  live <- readAOFM::read_aofm("tb", "issuance")
  utils::head(live, 3)
} else {
  cat("Live probe skipped; set READAOFM_RUN_LIVE_EXAMPLES=true to opt in.\n")
}
```
