---
title: "Table catalog and data sources"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Table catalog and data sources}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

`readAOFM` ships a local catalogue that maps AOFM Data Hub workbooks to search
terms, public selectors, parser families, and direct HTTPS source routes. This
article records the 23 parser-supported entries and the seven raw-only /
unsupported entries available through raw workbook download. Reader calls
retrieve current source workbooks; the catalogue itself supplies stable,
offline discovery metadata.

## The 23 parsed tables

The public selectors are exact. The table below is deliberately explicit so
that the table ID, family, and corresponding `read_aofm()` call are visible in
one place. The validation column is computed with `search_aofm()`, which is an
offline catalogue operation.

```{r parsed-catalog}
parsed_catalog <- data.frame(
  family = c(
    "End-of-financial-year summary",
    rep("End-of-month positions", 8),
    rep("Transactional issuance and buybacks", 7),
    rep("Syndication details", 2),
    rep("Ownership", 2),
    rep("Secondary-market turnover", 2),
    "Term premium"
  ),
  security = c(
    "summary",
    "aggregate", "aggregate", "tb", "tb", "tib", "tib", "tn", "tn",
    "tb", "tb", "tib", "tib", "tn", "retail", "slf",
    "tb", "tib",
    "ownership", "ownership",
    "tb", "tib",
    "termpremium"
  ),
  type = c(
    NA_character_,
    "dealt", "settlement", "dealt", "settlement", "dealt", "settlement",
    "dealt", "settlement",
    "issuance", "buyback", "issuance", "buyback", "issuance", NA, NA,
    "syndication", "syndication",
    "public", "nonresident",
    "turnover", "turnover",
       NA_character_
  ),
  id = c(
    "summary",
    "aggregate_position_dealt", "aggregate_position_settlement",
    "tb_position_dealt", "tb_position_settlement",
    "tib_position_dealt", "tib_position_settlement",
    "tn_position_dealt", "tn_position_settlement",
    "tb_issuance", "tb_buyback", "tib_issuance", "tib_buyback",
    "tn_issuance", "retail", "slf",
    "tb_syndication", "tib_syndication",
    "ownership_public", "ownership_nonresident",
    "tb_turnover", "tib_turnover",
    "termpremium"
  ),
  stringsAsFactors = FALSE
)

parsed_catalog$selector <- ifelse(
  is.na(parsed_catalog$type),
  sprintf('read_aofm("%s")', parsed_catalog$security),
  sprintf(
    'read_aofm("%s", "%s")',
    parsed_catalog$security,
    parsed_catalog$type
  )
)

parsed_catalog$catalog_match <- vapply(seq_len(nrow(parsed_catalog)), function(i) {
  query <- if (is.na(parsed_catalog$type[[i]])) {
    parsed_catalog$security[[i]]
  } else {
    paste(parsed_catalog$security[[i]], parsed_catalog$type[[i]])
  }
  parsed_catalog$id[[i]] %in% readAOFM::search_aofm(query)$id
}, logical(1))

stopifnot(nrow(parsed_catalog) == 23L, all(parsed_catalog$catalog_match))
parsed_catalog[, c("family", "id", "selector")]
```

The same catalogue can be searched with user-facing terms. These calls are
local and return metadata only because `read = FALSE` is the default:

```{r family-search}
readAOFM::search_aofm("treasury bond")[, c("id", "read_call")]
readAOFM::search_aofm("inflation")[, c("id", "read_call")]
readAOFM::search_aofm("secondary market")[, c("id", "read_call")]
```

## Raw-only workbook catalogue entries

The following seven catalogue IDs are available through the raw workbook
download path while the parser-supported workflow focuses on the 23 entries
above. They remain explicit catalogue rows with `supported = FALSE`, not
additional parser families:

```{r raw-only}
raw_only_ids <- c(
  "tb_issuance_conversion",
  "indexation_factors",
  "rmbs_transactions",
  "rmbs_auctions",
  "interest_rate_swaps",
  "cross_currency_swaps",
  "portfolio_overview"
)

raw_only <- data.frame(
  id = raw_only_ids,
  parser = NA_character_,
  access = "download_aofm_xlsx()",
  stringsAsFactors = FALSE
)

# This check reads package metadata only. It does not download a workbook;
# users should use search_aofm() rather than this internal object.
catalog_metadata <- get("aofm_index", envir = asNamespace("readAOFM"))
stopifnot(
  setequal(raw_only_ids, catalog_metadata$id[catalog_metadata$fn == "no function exists"]),
  all(catalog_metadata$fn[catalog_metadata$id %in% raw_only_ids] == "no function exists")
)
raw_only
```

The complete map is available without contacting AOFM:

```{r complete-catalog}
all_catalogue <- readAOFM::aofm_catalog(include_unsupported = TRUE)
stopifnot(
  nrow(all_catalogue) == 30L,
  sum(all_catalogue$supported) == 23L,
  all(grepl("^https://www\\.aofm\\.gov\\.au/sites/default/files/", all_catalogue$source_url)),
  all(!grepl("/media/", all_catalogue$source_url, fixed = TRUE))
)
all_catalogue[!all_catalogue$supported,
              c("table_id", "source_url", "file_name", "supported")]
```

The raw-only source filenames remain extension-compatible with their direct
routes, including the legacy `.xls` files. `portfolio_overview` is a verified
historical direct workbook that is not currently listed on the Data Hub; it is
retained as raw-only and is not presented as current or parser-supported.

`download_aofm_xlsx()` is the public raw-workbook helper. Its selectors are
the `security` and `type` values shown for parsed rows. The raw-only rows have
no parser and no public `read_aofm()` call; because their selector fields are
not populated, an unfiltered raw download is the available public route and
will also select other catalogue workbooks. The function writes files beneath
`data/` in the working directory and performs a live HTTPS request, so this
article does not run it automatically.

## Source access and scope

The public AOFM Data Hub provides these workbooks without a package username,
password, or API key. `search_aofm()` performs local discovery. Reader and raw
download calls request the workbook URL recorded in the package catalogue over
HTTPS, stage it in a temporary file (readers), or save it beneath `data/` (raw
downloads). Each call retrieves the current source rather than using a
persistent package cache. The opt-in `download_aofm_file()` API instead keeps a
verified, content-addressed workbook and metadata beneath the caller-selected
`.readAOFM/data/<table-id>/` root; `aofm_file_metadata()` and
`read_aofm_file()` inspect and parse local files without network access.

Readers and raw downloads accept `timeout`, `retries`, and `max_bytes` controls
(30 seconds, one retry, and 100 MiB by default). Managed downloads additionally
bound cache age, file count, total bytes, and lock wait time. These controls
keep network and filesystem work explicit and finite for scripts, tests, and
CRAN builds.

AOFM controls workbook URLs, sheet names, columns, and historical coverage.
Turnover observation periods and publication cadence are separate: current
turnover workbooks contain monthly observations, the Data Hub update is
quarterly with an approximately two-month lag, historical **By Tenor** is
monthly, and historical **By Category** is quarterly. The historical turnover
companions end in December 2025 and the redesigned current workbooks begin in
January 2026. `read_secondary()` joins them by `period`, `group`, and `name`,
uses current rows for overlaps, rejects duplicate keys, and attaches the two
source records as `aofm_sources`. The groups are `tenor`, `investor_type`,
`security`, `region`, and `counterparty`.

The continuity CSV retains `legacy_max = 2024-12-31` as the boundary of the
pinned retired-baseline comparison. The replacement historical workbook
extends through `2025-12-31`; together with the current `2026-01-31` onward
workbook it provides the documented no-gap transition. The packaged workbook
snapshots used by tests and vignettes are deterministic examples, not a
replacement for the live AOFM source or a runtime cache. A changed URL, HTML
error page, unavailable source, or changed workbook layout can therefore make
a read fail; report the table ID, selector, error text, and access date.
