---
title: "Table catalog and data sources"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Table catalog and data sources}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

`readAOFM` ships a local catalogue that maps AOFM Data Hub workbooks to search
terms, public selectors, and parser families. This article records the 23
parser-supported entries and the seven historical entries available through
raw workbook download. Reader calls retrieve current source workbooks; the
catalogue itself supplies stable discovery metadata.

## The 23 parsed tables

The public selectors are exact. The table below is deliberately explicit so
that the table ID, family, and corresponding `read_aofm()` call are visible in
one place. The validation column is computed with `search_aofm()`, which is an
offline catalogue operation.

```{r parsed-catalog}
parsed_catalog <- data.frame(
  family = c(
    "End-of-financial-year summary",
    rep("End-of-month positions", 8),
    rep("Transactional issuance and buybacks", 7),
    rep("Syndication details", 2),
    rep("Ownership", 2),
    rep("Secondary-market turnover", 2),
    "Term premium"
  ),
  security = c(
    "summary",
    "aggregate", "aggregate", "tb", "tb", "tib", "tib", "tn", "tn",
    "tb", "tb", "tib", "tib", "tn", "retail", "slf",
    "tb", "tib",
    "ownership", "ownership",
    "tb", "tib",
    "termpremium"
  ),
  type = c(
    NA_character_,
    "dealt", "settlement", "dealt", "settlement", "dealt", "settlement",
    "dealt", "settlement",
    "issuance", "buyback", "issuance", "buyback", "issuance", NA, NA,
    "syndication", "syndication",
    "public", "nonresident",
    "turnover", "turnover",
       NA_character_
  ),
  id = c(
    "summary",
    "aggregate_position_dealt", "aggregate_position_settlement",
    "tb_position_dealt", "tb_position_settlement",
    "tib_position_dealt", "tib_position_settlement",
    "tn_position_dealt", "tn_position_settlement",
    "tb_issuance", "tb_buyback", "tib_issuance", "tib_buyback",
    "tn_issuance", "retail", "slf",
    "tb_syndication", "tib_syndication",
    "ownership_public", "ownership_nonresident",
    "tb_turnover", "tib_turnover",
    "termpremium"
  ),
  stringsAsFactors = FALSE
)

parsed_catalog$selector <- ifelse(
  is.na(parsed_catalog$type),
  sprintf('read_aofm("%s")', parsed_catalog$security),
  sprintf(
    'read_aofm("%s", "%s")',
    parsed_catalog$security,
    parsed_catalog$type
  )
)

parsed_catalog$catalog_match <- vapply(seq_len(nrow(parsed_catalog)), function(i) {
  query <- if (is.na(parsed_catalog$type[[i]])) {
    parsed_catalog$security[[i]]
  } else {
    paste(parsed_catalog$security[[i]], parsed_catalog$type[[i]])
  }
  parsed_catalog$id[[i]] %in% readAOFM::search_aofm(query)$id
}, logical(1))

stopifnot(nrow(parsed_catalog) == 23L, all(parsed_catalog$catalog_match))
parsed_catalog[, c("family", "id", "selector")]
```

The same catalogue can be searched with user-facing terms. These calls are
local and return metadata only because `read = FALSE` is the default:

```{r family-search}
readAOFM::search_aofm("treasury bond")[, c("id", "read_call")]
readAOFM::search_aofm("inflation")[, c("id", "read_call")]
readAOFM::search_aofm("secondary market")[, c("id", "read_call")]
```

## Raw-workbook catalogue entries

The following seven catalogue IDs are available through the raw workbook
download path while the parser-supported workflow focuses on the 23 entries
above:

```{r raw-only}
raw_only_ids <- c(
  "tb_issuance_conversion",
  "indexation_factors",
  "rmbs_transactions",
  "rmbs_auctions",
  "interest_rate_swaps",
  "cross_currency_swaps",
  "portfolio_overview"
)

raw_only <- data.frame(
  id = raw_only_ids,
  parser = NA_character_,
  access = "download_aofm_xlsx()",
  stringsAsFactors = FALSE
)

# This check reads package metadata only. It does not download a workbook;
# users should use search_aofm() rather than this internal object.
catalog_metadata <- get("aofm_index", envir = asNamespace("readAOFM"))
stopifnot(
  setequal(raw_only_ids, catalog_metadata$id[catalog_metadata$fn == "no function exists"]),
  all(catalog_metadata$fn[catalog_metadata$id %in% raw_only_ids] == "no function exists")
)
raw_only
```

`download_aofm_xlsx()` is the public raw-workbook helper. Its selectors are
the `security` and `type` values shown for parsed rows. The raw-only rows have
no parser and no public `read_aofm()` call; because their selector fields are
not populated, an unfiltered raw download is the available public route and
will also select other catalogue workbooks. The function writes files beneath
`data/` in the working directory and performs a live HTTPS request, so this
article does not run it automatically.

## Source access and scope

The public AOFM Data Hub provides these workbooks without a package username,
password, or API key. `search_aofm()` performs local discovery. Reader and raw
download calls request the workbook URL recorded in the package catalogue over
HTTPS, stage it in a temporary file (readers), or save it beneath `data/` (raw
downloads). Each call retrieves the current source rather than using a
persistent package cache.

AOFM controls workbook URLs, publication cadence, sheet names, columns, and
historical coverage. For example, turnover is published quarterly with a
reporting lag, while several raw-only datasets are historical. A changed URL,
HTML error page, unavailable source, or changed workbook layout can therefore
make a read fail. Use the table ID, selector, error text, and access date when
reporting a source change. The packaged workbook snapshots used by tests and
vignettes are deterministic examples, not a replacement for the live AOFM
source or a runtime cache.
